---
title: "Ubuntu Commands"
author: "Tobias Romar"
date: "4 augusti 2017"
output: html_document
---

The purpose with document is describe some of the most common commands. 

#### 1. apt — Advanced Packaging Tool
Used to manage packages. Install, remove and upgrade any package.  

```{r eval = FALSE}
sudo apt install application-name
sudo apt-get remove application-name
sudo apt-get autoclean
``` 

#### 2. ls -  List
Find a file, or get a quick overview of what files exist in the current directory.

```{r eval = FALSE}
ls
ls -lh
ls -lS
``` 


#### 3. cd - Change Directory
Used to change and navigate through directories. Cd will assume you’re in your 
Home folder. Its use is straightforward: 

```{r eval = FALSE}
cd Pictures
cd cats/
cd -
cd ..
cd
``` 


#### 4. sudo - Super User…

You can’t do anything too dramatic to your system without it. sudo lets you 
run commands, install software, edit protected files, as the superuser.  

```{r eval = FALSE}
sudo edit /usr/share/applications/application.desktop
sudo apt-get install application-name
``` 

#### 5. cat - catenate

The cat command read data from files and outputs its content in the terminal. Using cat is the simplest way to display file contents at the command line.

```{r eval = FALSE}
cat examplefile.txt
``` 

cat - reads the text file so that you can see what is in it.
nano - opens the file so that you can see what is in it, with the option to add or delete some data.

#### 6. mkdir - Make directory 

Used to create a new folder or subfolder.
```{r eval = FALSE}
mkdir folderName
``` 

#### 7. cp - Copy-and-paste

First, you determine the file you want to copy and type the destination location to paste the file
```{r eval = FALSE}
cp src des
``` 

#### 8. rm - Remove

Command to remove your file or even your directory. 

```{r eval = FALSE}
rm myfile.txt
``` 

#### 9. grep - Find 

use the grep command to help finding the file based on given keywords.

```{r eval = FALSE}
grep user /etc/passwd
``` 

#### 10. pwd - Location

By typing to followingcommand you get the current location:

```{r eval = FALSE}
pwd
``` 



